Placeholder for article 3

### 2025-01-21 analysis/paper/paper.qmd
Pipe %>% replaced by |>

# Cleaning up DS and uniformising variable names
proto_tech_WIP <- proto_tech %>%
  tidyr::unite(col='ID', all_of(cols_to_concat_terrace), sep="-", remove=TRUE) %>% 
  rename(CLASS = Class, COREPREPPROD = CorePreparProd, PIECECOMPLETENESS = PieceCompleteness, CORTEX = Cortex, CORTEXLOCATION = CortexLocation, PLATFORMTYPE = PlatformType, PLATFORMCORTEX = PlatformCortex, SCARCOUNT = ScarCount, THICKNESS = Thickness,
        MAXWIDTH = MaxWidth, LENGTH = Length, COREPLATFORM = CorePlatform, MAINFACECORTEX = MainFaceCortex, 
        MAINFACESCARCOUNT = MainFaceScarCount, COREABANDON = CoreAbandon, RETOUCHEDPIECETYPOLOGY = RetouchedPieceTypology) %>% 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Right truncation on retouched piece" ~ "Truncation",
    RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed bladelet parcial" ~ "Backed bladelet",
    RETOUCHEDPIECETYPOLOGY == "Dentilhado" ~ "Denticulate",
    RETOUCHEDPIECETYPOLOGY == "Burin on truncation" ~ "Burin",
    TRUE ~ RETOUCHEDPIECETYPOLOGY)) %>% 
  dplyr::select(ID, CLASS, COREPREPPROD, PIECECOMPLETENESS, CORTEX, CORTEXLOCATION, PLATFORMTYPE, PLATFORMCORTEX,
                SCARCOUNT, THICKNESS, MAXWIDTH, LENGTH, COREPLATFORM, MAINFACECORTEX, MAINFACESCARCOUNT, COREABANDON, RETOUCHEDPIECETYPOLOGY)

gravettian_tech <- gravettian_tech %>% 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Splintered piece with notch" | RETOUCHEDPIECETYPOLOGY == "Splintered piece with lateral retouch" ~ "Splintered piece",
    RETOUCHEDPIECETYPOLOGY == "Retouched flake siret fragment" | RETOUCHEDPIECETYPOLOGY == "Retouched flake frag" ~ "Retouched flake fragment",
    RETOUCHEDPIECETYPOLOGY == "Retouched bladelet frag" ~ "Retouched bladelet fragment",
    RETOUCHEDPIECETYPOLOGY == "Multiple burin" | RETOUCHEDPIECETYPOLOGY == "Burin splintered piece" | RETOUCHEDPIECETYPOLOGY == "Burin on truncation" ~ "Burin",
    RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" ~ "Backed bladelet",
    TRUE ~ RETOUCHEDPIECETYPOLOGY)) %>% 
  mutate(CLASS = case_when(
    CLASS == "Flake" | CLASS == "ElongBlank" ~ "Blank",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    TRUE ~ CLASS))

@@

terrace_RM_WIP <- terrace_RM %>%
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "TYPE 2",
    RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "TYPE 3",
    RMTYPE == "TYPE 7" ~ "TYPE 4",
    RMTYPE == "TYPE 12" ~ "TYPE 5",
    RMTYPE == "TYPE 11" ~ "TYPE 6",
    RMTYPE == "TYPE 13" ~ "TYPE 7",
    RMTYPE == "TYPE 11F" ~ "TYPE 8",
    RMTYPE == "TYPE 13B" ~ "TYPE 9",
    RMTYPE == "Oolitic" ~ "TYPE 10",
    RMTYPE == "TYPE GB" ~ "TYPE 11",
    .default = RMTYPE)) %>% 
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%
  distinct(ID, .keep_all = TRUE)

context <- context %>%
  dplyr::rename("UNIT" = "Unit") %>% # Renaming first column from Unit to UNIT to use cols_to_concat
  unite(col='ID', cols_to_concat_terrace, sep="-", remove=FALSE) %>% # Unite UNIT with ID using cols_to_concat
  select(ID, Level, Spit, Code) %>% # Selecting important variables: ID refers to artifact ID; Level refers to archaeologycal layer; Spit refers to artificial levels within an archaeological layer; Code refers to type of artifact (e.g., LITHIC or BONE)
  distinct(ID, .keep_all = TRUE)

xyz <- xyz %>%
  unite(col='ID', cols_to_concat_terrace, sep="-", remove=FALSE) %>% # Unite UNIT with ID using cols_to_concat
  filter(Suffix == 0) %>% # To avoid several entries related to several points of the same artefact
  select(ID, X, Y, Z) %>% # Selecting important variables
  distinct(ID, .keep_all = TRUE)

@@

terrace_tech_WIP <- left_join(terrace_tech, field_data, by = "ID") # Join field and RM type tables
terrace_dataset <- full_join(terrace_tech_WIP, terrace_RM_WIP, by = "ID") # Base 12690
terrace_dataset <- terrace_dataset %>%
  filter(!Level %in% c("3", "4", "4C", "4D")) %>%  # Dropping NAs on the technological class drops dataset to 1651
  drop_na(RMTYPE) %>%  # Dropping NAs on the RM database drops dataset to 3635
  drop_na(CLASS) # Dropping NAs on the technological class drops dataset to 1647

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- terrace_dataset %>% 
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    Level == "7C" | Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7 to homogenise dataset
    TRUE ~ Level)) %>% # Adding technocomplex name based on the levels
  mutate(Technocomplex = case_when(
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian"))

@@

solutrean_WIP <- solutrean_JC %>% 
  filter(Layer %in% c("A", "B", "C")) %>% # Filtering Solutrean layers
  filter(Raw.material %in% c("Chert", "Chalcedony")) %>% # Filtering raw materials
  filter(Screen != "Crivo") %>%  # Filtering out all bucket lithics since they were not included in the RM analysis
  mutate_at("Date", str_replace, "20", "") %>% # Removing initial part of the year to match IDs
  dplyr::rename("YEAR" = "Date") %>% 
  tidyr::unite(col='ID', all_of(cols_to_concat_shelter), sep="-", remove=FALSE) # Concatenate YEAR and ID variables in one single variable (e.g. 05-500)

# Dataset filtering for chert/chalcedony and levels (A-C), 3298 + 61 solutrean points, n=3359
solutrean_original_chert <- solutrean_JC %>% 
  filter(Layer %in% c("A", "B", "C")) %>% # Filtering Solutrean layers
  filter(Raw.material %in% c("Chert", "Chalcedony"))

# In Cascalheira 2010 Retouched piece was not a category but a variable. To match the other DBs, this category was added
solutrean_WIP <- solutrean_WIP %>% 
  mutate(Class = case_when( # This alteration is being done on the original class column
    Retouched == "Yes" ~ "Retouched piece", # Whenever retouch is present, class is now Retouched piece
    TRUE ~ Class)) %>% 
  mutate(CLASS = case_when( # Creating a column for clustered classes, to match the other DBs
    Class == "Blade" | Class == "Bladelet" ~ "ElongBlank", # Cluster blades with bladelets. The distinction can come later through a mutate and the measurements
    Class == "Core front" | Class == "Core tablet" | Class == "Core trimming elem." | Class == "Crested piece" ~ "CorePrepProd", # Cluster all preparation products
    Class == "Fragment" ~ "Shatter",
    Class == "Flake distal frag." | Class == "Flake mesial frag." | Class == "Flake prox. frag." ~ "FlakeFrag", # Cluster flake frags into a single category
    Class == "Blade distal frag." | Class == "Blade mesial frag." | Class == "Blade prox. frag." 
    | Class == "Bladelet distal frag." | Class == "Bladelet mesial frag." | Class == "Bladelet prox. frag." ~ "ElongBlankFrag", # Cluster elong frags into a single category
    TRUE ~ Class)) %>% 
  mutate(COREPREPPROD = case_when( # Creating a column for the types of core preparation and maintenance products (COREPREPPROD)
    Class == "Core front" ~ "Core front",
    Class == "Core tablet" ~ "Core tablet",
    Class == "Core trimming elem." ~ "Core trim",
    Class == "Crested piece" ~ "Crested piece")) %>% 
  mutate(PIECECOMPLETENESS = case_when( # Creating a column for the portions present in flake or elongblank fragments to match other DBs
    Class == "Flake distal frag." | Class == "Blade distal frag." | Class == "Bladelet distal frag." ~ "Distal",
    Class == "Flake mesial frag." | Class == "Blade mesial frag." | Class == "Bladelet mesial frag." ~ "Mesial",
    Class == "Flake prox. frag." | Class == "Blade prox. frag." | Class == "Bladelet prox. frag." ~ "Proximal")) %>% 
  rename(PLATFORMTYPE = Platform_type) %>% # Renaming column
  mutate(PLATFORMTYPE = case_when( # Renaming variables to match correspondent variables in the other DBs
    PLATFORMTYPE == "Multifacetted" ~ "Faceted",
    PLATFORMTYPE == "Unfacetted" ~ "Plain",
    TRUE ~ PLATFORMTYPE)) %>%
  mutate(Type = case_when( # Translating retouched types from Zilhão 1997 to simplified typologies
    Type == "92b" ~ "Retouched piece fragment",
    Type == "92a" ~ "Retouched blank",
    Type == "76" ~ "Splintered piece",
    Type == "75" ~ "Denticulate",
    Type == "74" ~ "Notch",
    Type == "66" ~ "Retouched blade",
    Type == "61" ~ "Truncation",
    Type == "28" ~ "Burin",
    Type == "7" | Type == "5b" | Type == "4"  ~ "Endscrapper",
    TRUE ~ Type)) %>%
  rename(LEVEL = Layer, UNIT = Unit, CORTEX = Percent_of_cortex, CORTEXLOCATION = Loc_of_cortex, THICKNESS = Thickness,
         MAXWIDTH = Width, LENGTH = Length, COREPLATFORM = Core_platfform, SCARCOUNT = N_of_Scars, COREABANDON = Core_abandon,
         RETOUCHEDPIECETYPOLOGY = Type)  # Renaming column

solutrean_tech <- solutrean_WIP %>%
  dplyr::select(ID, CLASS, LEVEL, UNIT, COREPREPPROD, PIECECOMPLETENESS, CORTEX, CORTEXLOCATION, PLATFORMTYPE,
                SCARCOUNT, THICKNESS, MAXWIDTH, LENGTH, COREPLATFORM, COREABANDON, RETOUCHEDPIECETYPOLOGY)

solutrean_points <- solutrean_points %>% 
  mutate_at("YEAR", str_replace, "20", "") %>% # Removing initial part of the year to match IDs
  tidyr::unite(col='ID', all_of(cols_to_concat_shelter), sep="-", remove=TRUE)

@@

solutrean_tech <- solutrean_tech %>% # Originally 2228
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets (down to 1993)

shelter_RM_WIP <- shelter_RM %>%
  mutate_at("ID", str_replace, "VB", "") %>% # Remove "VB" from ID
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "TYPE 2",
    RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "TYPE 3",
    RMTYPE == "TYPE 7" ~ "TYPE 4",
    RMTYPE == "TYPE 12" ~ "TYPE 5",
    RMTYPE == "TYPE 11" ~ "TYPE 6",
    RMTYPE == "TYPE 13" ~ "TYPE 7",
    RMTYPE == "TYPE 11F" ~ "TYPE 8",
    RMTYPE == "TYPE 13B" ~ "TYPE 9",
    RMTYPE == "Oolitic" ~ "TYPE 10",
    RMTYPE == "TYPE GB" ~ "TYPE 11",
    RMTYPE == "ND_A" ~ "INDET",
    .default = RMTYPE)) %>%
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%  # Select variables
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets

solutrean_dataset <- full_join(solutrean_tech, shelter_RM_WIP, by = "ID") # The complete dataset has 2818 (44 points) entries
solutrean_dataset <- solutrean_dataset %>%
  drop_na(RMTYPE) %>%   # Dropping NAs on the RM database drops dataset to 1514
  drop_na(CLASS) # Dropping NAs on the technological class drops dataset to 914

@@

# Solutrean statistics
# DS preparation
solutrean_stats <- solutrean_dataset %>% 
  filter(RMTYPE != "INDET") %>%
   mutate(CLASS = case_when(
    CLASS == "Thinning flake" ~ "ThinningFlake",
    CLASS == "ElongBlank" ~ "ElongatedProd",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    CLASS == "Retouched piece" ~ "RetouchedPiece",
    TRUE ~ CLASS)) %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "BlankFragProx", # To add proximal fragments to ratio
    TRUE ~ CLASS))

@@

# Boxplot for individual measurements
# Weight
solutrean_stats_weight <- solutrean_stats %>%
  filter(CLASS %in% c("Flake", "ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, WEIGHT, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(x="Raw material types" ,y="Weight (gr)") +
  coord_cartesian(ylim =  c(0, 50)) +
  scale_fill_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6", "#BDC7C9")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Proto-Solutrean statistics
# DS prep
proto_stats <- phase_dataset %>% 
  filter(Level %in% c("4E", "4E/5")) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "BlankFragProx", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>%
  mutate(elong = MAXWIDTH*2-LENGTH) %>%
  mutate(CLASS = case_when(
    CLASS == "Blank" & elong > 0 ~ "Flake", # QUESTION: do I keep flakes and elong separate or join into blanks?
    CLASS == "Blank" & elong <= 0 ~ "ElongatedProd",
    TRUE ~ CLASS))

@@

# Boxplots weight
proto_stats_weight <- proto_stats %>%
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) %>% 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","GB","TL")) %>%
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10"))) %>%
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) %>%
  ggplot(aes(RMTYPE, WEIGHT, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(x="Raw material types", y="Weight (gr)") +
  coord_cartesian(ylim =  c(0, 50)) +
  scale_fill_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

>>>>

# Cleaning up DS and uniformising variable names
proto_tech_WIP <- proto_tech |>
  tidyr::unite(col='ID', all_of(cols_to_concat_terrace), sep="-", remove=TRUE) |> 
  rename(CLASS = Class, COREPREPPROD = CorePreparProd, PIECECOMPLETENESS = PieceCompleteness, CORTEX = Cortex, CORTEXLOCATION = CortexLocation, PLATFORMTYPE = PlatformType, PLATFORMCORTEX = PlatformCortex, SCARCOUNT = ScarCount, THICKNESS = Thickness,
        MAXWIDTH = MaxWidth, LENGTH = Length, COREPLATFORM = CorePlatform, MAINFACECORTEX = MainFaceCortex, 
        MAINFACESCARCOUNT = MainFaceScarCount, COREABANDON = CoreAbandon, RETOUCHEDPIECETYPOLOGY = RetouchedPieceTypology) |> 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Right truncation on retouched piece" ~ "Truncation",
    RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed bladelet parcial" ~ "Backed bladelet",
    RETOUCHEDPIECETYPOLOGY == "Dentilhado" ~ "Denticulate",
    RETOUCHEDPIECETYPOLOGY == "Burin on truncation" ~ "Burin",
    TRUE ~ RETOUCHEDPIECETYPOLOGY)) |> 
  dplyr::select(ID, CLASS, COREPREPPROD, PIECECOMPLETENESS, CORTEX, CORTEXLOCATION, PLATFORMTYPE, PLATFORMCORTEX,
                SCARCOUNT, THICKNESS, MAXWIDTH, LENGTH, COREPLATFORM, MAINFACECORTEX, MAINFACESCARCOUNT, COREABANDON, RETOUCHEDPIECETYPOLOGY)

gravettian_tech <- gravettian_tech |> 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Splintered piece with notch" | RETOUCHEDPIECETYPOLOGY == "Splintered piece with lateral retouch" ~ "Splintered piece",
    RETOUCHEDPIECETYPOLOGY == "Retouched flake siret fragment" | RETOUCHEDPIECETYPOLOGY == "Retouched flake frag" ~ "Retouched flake fragment",
    RETOUCHEDPIECETYPOLOGY == "Retouched bladelet frag" ~ "Retouched bladelet fragment",
    RETOUCHEDPIECETYPOLOGY == "Multiple burin" | RETOUCHEDPIECETYPOLOGY == "Burin splintered piece" | RETOUCHEDPIECETYPOLOGY == "Burin on truncation" ~ "Burin",
    RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" ~ "Backed bladelet",
    TRUE ~ RETOUCHEDPIECETYPOLOGY)) |> 
  mutate(CLASS = case_when(
    CLASS == "Flake" | CLASS == "ElongBlank" ~ "Blank",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    TRUE ~ CLASS))

@@

terrace_RM_WIP <- terrace_RM |>
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "TYPE 2",
    RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "TYPE 3",
    RMTYPE == "TYPE 7" ~ "TYPE 4",
    RMTYPE == "TYPE 12" ~ "TYPE 5",
    RMTYPE == "TYPE 11" ~ "TYPE 6",
    RMTYPE == "TYPE 13" ~ "TYPE 7",
    RMTYPE == "TYPE 11F" ~ "TYPE 8",
    RMTYPE == "TYPE 13B" ~ "TYPE 9",
    RMTYPE == "Oolitic" ~ "TYPE 10",
    RMTYPE == "TYPE GB" ~ "TYPE 11",
    .default = RMTYPE)) |> 
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) |>
  distinct(ID, .keep_all = TRUE)

context <- context |>
  dplyr::rename("UNIT" = "Unit") |> # Renaming first column from Unit to UNIT to use cols_to_concat
  unite(col='ID', cols_to_concat_terrace, sep="-", remove=FALSE) |> # Unite UNIT with ID using cols_to_concat
  select(ID, Level, Spit, Code) |> # Selecting important variables: ID refers to artifact ID; Level refers to archaeologycal layer; Spit refers to artificial levels within an archaeological layer; Code refers to type of artifact (e.g., LITHIC or BONE)
  distinct(ID, .keep_all = TRUE)

xyz <- xyz |>
  unite(col='ID', cols_to_concat_terrace, sep="-", remove=FALSE) |> # Unite UNIT with ID using cols_to_concat
  filter(Suffix == 0) |> # To avoid several entries related to several points of the same artefact
  select(ID, X, Y, Z) |> # Selecting important variables
  distinct(ID, .keep_all = TRUE)

@@

terrace_tech_WIP <- left_join(terrace_tech, field_data, by = "ID") # Join field and RM type tables
terrace_dataset <- full_join(terrace_tech_WIP, terrace_RM_WIP, by = "ID") # Base 12690
terrace_dataset <- terrace_dataset |>
  filter(!Level %in% c("3", "4", "4C", "4D")) |>  # Dropping NAs on the technological class drops dataset to 1651
  drop_na(RMTYPE) |>  # Dropping NAs on the RM database drops dataset to 3635
  drop_na(CLASS) # Dropping NAs on the technological class drops dataset to 1647

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- terrace_dataset |> 
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    Level == "7C" | Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7 to homogenise dataset
    TRUE ~ Level)) |> # Adding technocomplex name based on the levels
  mutate(Technocomplex = case_when(
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian"))

@@

solutrean_WIP <- solutrean_JC |> 
  filter(Layer %in% c("A", "B", "C")) |> # Filtering Solutrean layers
  filter(Raw.material %in% c("Chert", "Chalcedony")) |> # Filtering raw materials
  filter(Screen != "Crivo") |>  # Filtering out all bucket lithics since they were not included in the RM analysis
  mutate_at("Date", str_replace, "20", "") |> # Removing initial part of the year to match IDs
  dplyr::rename("YEAR" = "Date") |> 
  tidyr::unite(col='ID', all_of(cols_to_concat_shelter), sep="-", remove=FALSE) # Concatenate YEAR and ID variables in one single variable (e.g. 05-500)

# Dataset filtering for chert/chalcedony and levels (A-C), 3298 + 61 solutrean points, n=3359
solutrean_original_chert <- solutrean_JC |> 
  filter(Layer %in% c("A", "B", "C")) |> # Filtering Solutrean layers
  filter(Raw.material %in% c("Chert", "Chalcedony"))

# In Cascalheira 2010 Retouched piece was not a category but a variable. To match the other DBs, this category was added
solutrean_WIP <- solutrean_WIP |> 
  mutate(Class = case_when( # This alteration is being done on the original class column
    Retouched == "Yes" ~ "Retouched piece", # Whenever retouch is present, class is now Retouched piece
    TRUE ~ Class)) |> 
  mutate(CLASS = case_when( # Creating a column for clustered classes, to match the other DBs
    Class == "Blade" | Class == "Bladelet" ~ "ElongBlank", # Cluster blades with bladelets. The distinction can come later through a mutate and the measurements
    Class == "Core front" | Class == "Core tablet" | Class == "Core trimming elem." | Class == "Crested piece" ~ "CorePrepProd", # Cluster all preparation products
    Class == "Fragment" ~ "Shatter",
    Class == "Flake distal frag." | Class == "Flake mesial frag." | Class == "Flake prox. frag." ~ "FlakeFrag", # Cluster flake frags into a single category
    Class == "Blade distal frag." | Class == "Blade mesial frag." | Class == "Blade prox. frag." 
    | Class == "Bladelet distal frag." | Class == "Bladelet mesial frag." | Class == "Bladelet prox. frag." ~ "ElongBlankFrag", # Cluster elong frags into a single category
    TRUE ~ Class)) |> 
  mutate(COREPREPPROD = case_when( # Creating a column for the types of core preparation and maintenance products (COREPREPPROD)
    Class == "Core front" ~ "Core front",
    Class == "Core tablet" ~ "Core tablet",
    Class == "Core trimming elem." ~ "Core trim",
    Class == "Crested piece" ~ "Crested piece")) |> 
  mutate(PIECECOMPLETENESS = case_when( # Creating a column for the portions present in flake or elongblank fragments to match other DBs
    Class == "Flake distal frag." | Class == "Blade distal frag." | Class == "Bladelet distal frag." ~ "Distal",
    Class == "Flake mesial frag." | Class == "Blade mesial frag." | Class == "Bladelet mesial frag." ~ "Mesial",
    Class == "Flake prox. frag." | Class == "Blade prox. frag." | Class == "Bladelet prox. frag." ~ "Proximal")) |> 
  rename(PLATFORMTYPE = Platform_type) |> # Renaming column
  mutate(PLATFORMTYPE = case_when( # Renaming variables to match correspondent variables in the other DBs
    PLATFORMTYPE == "Multifacetted" ~ "Faceted",
    PLATFORMTYPE == "Unfacetted" ~ "Plain",
    TRUE ~ PLATFORMTYPE)) |>
  mutate(Type = case_when( # Translating retouched types from Zilhão 1997 to simplified typologies
    Type == "92b" ~ "Retouched piece fragment",
    Type == "92a" ~ "Retouched blank",
    Type == "76" ~ "Splintered piece",
    Type == "75" ~ "Denticulate",
    Type == "74" ~ "Notch",
    Type == "66" ~ "Retouched blade",
    Type == "61" ~ "Truncation",
    Type == "28" ~ "Burin",
    Type == "7" | Type == "5b" | Type == "4"  ~ "Endscrapper",
    TRUE ~ Type)) |>
  rename(LEVEL = Layer, UNIT = Unit, CORTEX = Percent_of_cortex, CORTEXLOCATION = Loc_of_cortex, THICKNESS = Thickness,
         MAXWIDTH = Width, LENGTH = Length, COREPLATFORM = Core_platfform, SCARCOUNT = N_of_Scars, COREABANDON = Core_abandon,
         RETOUCHEDPIECETYPOLOGY = Type)  # Renaming column

solutrean_tech <- solutrean_WIP |>
  dplyr::select(ID, CLASS, LEVEL, UNIT, COREPREPPROD, PIECECOMPLETENESS, CORTEX, CORTEXLOCATION, PLATFORMTYPE,
                SCARCOUNT, THICKNESS, MAXWIDTH, LENGTH, COREPLATFORM, COREABANDON, RETOUCHEDPIECETYPOLOGY)

solutrean_points <- solutrean_points |> 
  mutate_at("YEAR", str_replace, "20", "") |> # Removing initial part of the year to match IDs
  tidyr::unite(col='ID', all_of(cols_to_concat_shelter), sep="-", remove=TRUE)

@@

solutrean_tech <- solutrean_tech |> # Originally 2228
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets (down to 1993)

shelter_RM_WIP <- shelter_RM |>
  mutate_at("ID", str_replace, "VB", "") |> # Remove "VB" from ID
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "TYPE 2",
    RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "TYPE 3",
    RMTYPE == "TYPE 7" ~ "TYPE 4",
    RMTYPE == "TYPE 12" ~ "TYPE 5",
    RMTYPE == "TYPE 11" ~ "TYPE 6",
    RMTYPE == "TYPE 13" ~ "TYPE 7",
    RMTYPE == "TYPE 11F" ~ "TYPE 8",
    RMTYPE == "TYPE 13B" ~ "TYPE 9",
    RMTYPE == "Oolitic" ~ "TYPE 10",
    RMTYPE == "TYPE GB" ~ "TYPE 11",
    RMTYPE == "ND_A" ~ "INDET",
    .default = RMTYPE)) |>
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) |>  # Select variables
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets

solutrean_dataset <- full_join(solutrean_tech, shelter_RM_WIP, by = "ID") # The complete dataset has 2818 (44 points) entries
solutrean_dataset <- solutrean_dataset |>
  drop_na(RMTYPE) |>   # Dropping NAs on the RM database drops dataset to 1514
  drop_na(CLASS) # Dropping NAs on the technological class drops dataset to 914

@@

# Solutrean statistics
# DS preparation
solutrean_stats <- solutrean_dataset |> 
  filter(RMTYPE != "INDET") |>
   mutate(CLASS = case_when(
    CLASS == "Thinning flake" ~ "ThinningFlake",
    CLASS == "ElongBlank" ~ "ElongatedProd",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    CLASS == "Retouched piece" ~ "RetouchedPiece",
    TRUE ~ CLASS)) |> 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "BlankFragProx", # To add proximal fragments to ratio
    TRUE ~ CLASS))

@@

# Boxplot for individual measurements
# Weight
solutrean_stats_weight <- solutrean_stats |>
  filter(CLASS %in% c("Flake", "ElongatedProd", "Core", "RetouchedPiece")) |> 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","GB","TL")) |>
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10", "TYPE 11", "TL"))) |>
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) |>
  ggplot(aes(RMTYPE, WEIGHT, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(x="Raw material types" ,y="Weight (gr)") +
  coord_cartesian(ylim =  c(0, 50)) +
  scale_fill_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6", "#BDC7C9")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Proto-Solutrean statistics
# DS prep
proto_stats <- phase_dataset |> 
  filter(Level %in% c("4E", "4E/5")) |> 
  filter(RMTYPE != "INDET") |> 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "BlankFragProx", # To add proximal fragments to ratio
    TRUE ~ CLASS)) |>
  mutate(elong = MAXWIDTH*2-LENGTH) |>
  mutate(CLASS = case_when(
    CLASS == "Blank" & elong > 0 ~ "Flake", # QUESTION: do I keep flakes and elong separate or join into blanks?
    CLASS == "Blank" & elong <= 0 ~ "ElongatedProd",
    TRUE ~ CLASS))

@@

# Boxplots weight
proto_stats_weight <- proto_stats |>
  filter(CLASS %in% c("Flake","ElongatedProd", "Core", "RetouchedPiece")) |> 
  filter(WEIGHT !="0" & !RMTYPE %in% c("INDET","GB","TL")) |>
  mutate(RMTYPE = factor(RMTYPE, levels = c("TYPE 1","TYPE 2", "TYPE 3", "TYPE 4",
                                        "TYPE 5", "TYPE 6","TYPE 7", 
                                        "TYPE 8", "TYPE 9", "TYPE 10"))) |>
  mutate(CLASS = factor(CLASS, levels = c("Flake", "ElongatedProd", "Core", "RetouchedPiece"))) |>
  ggplot(aes(RMTYPE, WEIGHT, fill = RMTYPE)) +
  geom_boxplot(aes(fill=factor(RMTYPE)), outlier.shape=21) +
  facet_wrap(~CLASS) +
  stat_summary(fun.y=mean, geom="point", shape=23) +
  labs(x="Raw material types", y="Weight (gr)") +
  coord_cartesian(ylim =  c(0, 50)) +
  scale_fill_manual(values = c("#8B7B8B", "#FFE1FF", "#EED2EE", "#D8BFD8",
                              "#CDB5CD", "#2B4F60" , "#4A708B", "#798777",
                            "#9FB6CD", "#BDD2B6")) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
